---
title: "PostgreSQL Tutorial - Home - Section 2"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
---

# Querying Data

## SELECT

`SELECT` statement template:
```
SELECT
  column_1,
  column_2,
  ...
FROM
  table_name;
```

Before we do a query, let's first connect to the _**dvdrental**_ database.

```{r}
library(DBI)
library(odbc)
library(RPostgres)
```

```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

```{sql connection=con_dvd}
SELECT
  *
FROM
  customer;
```

```{sql connection=con_dvd}
SELECT
  first_name,
  last_name,
  email
FROM
  customer;
```


## ORDER BY

`ORDER BY` clause template:
```
SELECT
  column_1,
  column_2
FROM
  tbl_name
ORDER BY
  column_1 ASC,
  column_2 DESC;
```

```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
ORDER BY
  first_name ASC;
```

```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
ORDER BY
  last_name DESC;
```

```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
ORDER BY
  first_name ASC,
  last_name DESC;
```


## SELECT DISTINCT

`SELECT DISTINCT` template:
```
SELECT
  DISTINCT column_1
FROM
  table_name;
```

If you specify multiple columns, the `DISTINCT` clause will evaluate the duplicate based on the combination of values of these columns.

In the following example, both `column_1` and `column_2` will be evaluated for duplicates.

```
SELECT
  DISTINCT column_1,
  column_2
FROM
  table_name;
```

PostgreSQL also provides the `DISTINCT ON (expression)` clause to keep the "first" row of each group of duplicates.

```
SELECT
  DISTINCT ON (column_1) column_1_alias
  column_2
FROM
  table_name;
ORDER BY
  column_1,
  column_2;
```

Note that the expression in parentheses following `DISTINCT ON` must be the first (or left-most) expression following the `ORDER BY` clause.

### Examples

First, let's create a table called `t1` in the `example` database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{sql connection=con_ex}
DROP TABLE IF EXISTS t1;
```


```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS t1(
  id      serial   NOT NULL PRIMARY KEY,
  fcolor  varchar, 
  bcolor  varchar  
);
```

```{sql connection=con_ex}
INSERT INTO t1 (bcolor, fcolor) 
VALUES 
  ('red',   'red'),
  ('red',   'red'),
  ('red',   NULL),
  (NULL,    'red'),
  ('red',   'green'),
  ('red',   'blue'),
  ('green', 'red'),
  ('green', 'blue'),
  ('green', 'green'),
  ('blue',  'red'),
  ('blue',  'green'),
  ('blue',  'blue');
```

```{sql connection=con_ex}
SELECT
  id,
  bcolor,
  fcolor
FROM
  t1;
```

#### `DISTINCT` on 1 column

```{sql connection=con_ex}
SELECT
  DISTINCT bcolor
FROM
  t1
ORDER BY
  bcolor;
```

#### `DISTINCT on multiple columns

```{sql connection=con_ex}
SELECT
  DISTINCT bcolor,
  fcolor
FROM
  t1
ORDER BY
  bcolor,
  fcolor;
```

#### `DISTINCT ON ... ORDER BY` example

```{sql connection=con_ex}
SELECT
  DISTINCT ON 
  (bcolor) bcolor,
  fcolor
FROM
  t1
ORDER BY
  bcolor,
  fcolor;
```

Disconnect from _**example**_ database.
```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex); }
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```