---
title: "PostgreSQL Tutorial - Home - Section 13"
output: 
  html_notebook:
    theme: sandstone
    highlight: zenburn
    toc: true
    toc_float: true
---

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```

Helper function for simplifying database disconnection/removal.

```{r}
my_dbDisconnect <- function(x) {
  x_expr <- rlang::enexpr(x)
  if (exists(rlang::as_string(x_expr))) { dbDisconnect(x); rm(x); }
}
```

# Understanding PostgreSQL constraints

## PRIMARY KEY

There are a few different ways to define primary keys when creating a table.

Single column primary key.

```
CREATE TABLE table_name(
  col_1  INT  PRIMARY KEY,
  col_2  VARCHAR,
  col_3  VARCHAR,
  ...
);
```

Single column primary key (another way).

```
CREATE TABLE table_name(
  col_1 INT,
  col_2 VARCHAR,
  col_3 VARCHAR,
  ...
  PRIMARY KEY(col_1)
);
```

Multiple column primary key.

```
CREATE TABLE table_name(
  col_1 INT,
  col_2 VARCHAR,
  col_3 VARCHAR,
  ...
  PRIMARY KEY(col_1, col_2)
);
```

Connect to _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

We'll create some new tables related to purchase orders to demonstrate, "po_headers" and "po_items".

```{sql connection=con_ex}
CREATE TABLE po_headers(
  po_no             INTEGER  PRIMARY KEY,
  vendor_no         INTEGER,
  description       TEXT,
  shipping_address  TEXT
);
```

```{sql connection=con_ex}
CREATE TABLE po_items(
  po_no       INTEGER,
  item_no     INTEGER,
  product_no  INTEGER,
  qty         INTEGER,
  net_price   NUMERIC,
  PRIMARY KEY(po_no, item_no)
);
```

If you don’t specify explicitly the name for primary key constraint, PostgreSQL will assign a default name to the primary key constraint. By default, PostgreSQL uses `table_name_pkey` as the default name for the primary key constraint. In this example, PostgreSQL creates the primary key constraint with the name `po_items_pkey` for the "po_items" table.

To drop a primary key constraint, use `ALTER TABLE` with `DROP CONSTRAINT`.

```{sql connection=con_ex}
ALTER TABLE
  po_items
DROP CONSTRAINT
  po_items_pkey;
```

To add a primary key constraint to an already existing table, use `ALTER TABLE` with `ADD PRIMARY KEY`.

```{sql connection=con_ex}
ALTER TABLE
  po_items
ADD 
  PRIMARY KEY(po_no, item_no);
```

```{r}
my_dbDisconnect(con_ex)
```



```{r echo=FALSE}
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
