---
title: "PostgreSQL Tutorial - Home - Section 2"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
---

# Filtering Data

## WHERE

`WHERE` template:
```
SELECT
  column_1,
  column_2,
  ...
FROM
  table_name
WHERE
  conditions;
```

| Operator      | Description           |
|:-------------:|:----------------------|
| `=`           | Equal                 |
| `>`           | Greater than          |
| `<`           | Less than             |
| `>=`          | Greater than or equal |
| `<=`          | Less than or equal    |
| `<>` or `!=`  | Not equal             |
| `AND`         | Boolean AND           |
| `OR`          | Boolean OR            |

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```

Connect to _**dvdrental**_ database.
```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

Some `WHERE` clause examples:

```{sql connection=con_dvd}
SELECT
  last_name,
  first_name
FROM
  customer
WHERE
  first_name = 'Jamie';
```

`AND` operator:

```{sql connection=con_dvd}
SELECT
  last_name,
  first_name
FROM
  customer
WHERE
  first_name = 'Jamie'
  AND
  last_name = 'Rice';
```

`OR` operator:

```{sql connection=con_dvd}
SELECT
  customer_id,
  amount,
  payment_date
FROM
  payment
WHERE
  amount <= 1
  OR
  amount >= 8;
```


## LIMIT and OFFSET

`LIMIT` template:
```
SELECT
  col_1,
  col_2
FROM
  table_name
LIMIT n;
```

`OFFSET` template:
```
SELECT
  col_1,
  col_2
FROM
  table_name
LIMIT n OFFSET m; -- skip m rows of result before returning n records
```

`LIMIT` example:
```{sql connection=con_dvd}
SELECT
  film_id,
  title,
  release_year
FROM
  film
ORDER BY
  film_id
LIMIT 5;
```

`LIMIT ... OFFSET` example:
```{sql connection=con_dvd}
SELECT
  film_id,
  title,
  release_year
FROM
  film
ORDER BY
  film_id
LIMIT 5 OFFSET 3;
```

`LIMIT` is helpful for getting the highest and lowest values in a column.

```{sql connection=con_dvd}
SELECT
  film_id,
  title,
  rental_rate
FROM
  film
ORDER BY
  rental_rate ASC
LIMIT 10;
```

```{sql connection=con_dvd}
SELECT
  film_id,
  title,
  rental_rate
FROM
  film
ORDER BY
  rental_rate DESC
LIMIT 10;
```


## FETCH

`LIMIT` is not part of the SQL standard, but `FETCH` is (as of SQL:2008).

Here are some templates:

```
SELECT
  col_1
  col_2
FROM
  table_name
ORDER BY
  col_1
OFFSET m ROWS
FETCH FIRST n ROWS ONLY;
```

```
SELECT
  col_1
  col_2
FROM
  table_name
ORDER BY
  col_1
OFFSET m ROW
FETCH NEXT n ROW ONLY;
```

`FETCH` examples:
```{sql connection=con_dvd}
SELECT
  film_id,
  title
FROM
  film
ORDER BY
  title
                      -- If OFFSET not defined, default is "OFFSET 0 ROWS"
FETCH FIRST ROW ONLY; -- if n not defined, default is n=1... see ex. below
```

```{sql connection=con_dvd}
SELECT
  film_id,
  title
FROM
  film
ORDER BY
  title
OFFSET 0 ROWS
FETCH FIRST 1 ROW ONLY;
```

```{sql connection=con_dvd}
SELECT
  film_id,
  title
FROM
  film
ORDER BY
  title
-- OFFSET 0 ROWS
FETCH FIRST 5 ROWS ONLY;
```

```{sql connection=con_dvd}
SELECT
  film_id,
  title
FROM
  film
ORDER BY
  title
OFFSET 5 ROWS
FETCH FIRST 5 ROWS ONLY;
```


## IN

`IN` operator template:
```
SELECT
  column_1,
  column_2
FROM
  table_name
WHERE
  column_1 IN (value_1, value_2, value_3);
```

`IN` operator with subquery:
```
SELECT
  column_1,
  column_2
FROM
  table_1_name
WHERE
  column_1 IN (SLECT column_a FROM table_2_name);
```

`IN` operator example:
```{sql connection=con_dvd}
SELECT
  customer_id,
  rental_id,
  return_date
FROM
  rental
WHERE
  customer_id IN (1, 2)
ORDER BY
  return_date DESC;
```

`NOT IN` operator example:
```{sql connection=con_dvd}
SELECT
  customer_id,
  rental_id
  return_date
FROM
  rental
WHERE
  customer_id NOT IN (1, 2);
```

`IN` with a subquery example:

First, let's compose the subquery.
```{sql connection=con_dvd}
SELECT
  customer_id
FROM
  rental
WHERE
 CAST (return_date AS DATE) = '2005-05-27';
```

Then, let's use the subquery in a larger `IN`-operator query.
```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
WHERE
  customer_id IN (
    SELECT
      customer_id
    FROM
      rental
    WHERE
      CAST (return_date AS DATE) = '2005-05-27'
  );
```


## BETWEEN

`BETWEEN` template:
```
SELECT
  column_1,
  column_2
FROM
  table_name
WHERE
  column_1 BETWEEN low AND high;
```

`BETWEEN` example:
```{sql connection=con_dvd}
SELECT
  customer_id,
  payment_id,
  amount
FROM
  payment
WHERE
  amount BETWEEN 8 AND 9
ORDER BY
  amount ASC;
```

`NOT BETWEEN` example:
```{sql connection=con_dvd}
SELECT
  customer_id,
  payment_id,
  amount
FROM
  payment
WHERE 
  amount NOT BETWEEN 8 AND 9
ORDER BY
  amount DESC;
```

```{sql connection=con_dvd}
SELECT
  customer_id,
  payment_id,
  amount,
  payment_date
FROM
  payment
WHERE
  payment_date BETWEEN '2007-02-07' AND '2007-02-15';
```


## LIKE

`LIKE` operator template:
```
SELECT
  column_1,
  column_2
FROM
  table_name
WHERE
  column_1 LIKE 'A%';
```

`LIKE` operator examples:
```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
WHERE
  first_name LIKE 'Jen%';
```

```{sql connection=con_dvd}
SELECT
  'foo' LIKE 'foo' AS ex1,
  'foo' LIKE 'f%'  AS ex2,
  'foo' LIKE '_o_' AS ex3,
  'bar' LIKE 'b_'  AS ex4;
```

```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
WHERE
  first_name LIKE '_her%';
```

`NOT LIKE` example:
```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
WHERE
  first_name NOT LIKE 'Jen%';
```

## PostgreSQL's extensions of the `LIKE` operator

### ILIKE

`ILIKE` is just the case *I*nsenstive version of `LIKE`.

```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
WHERE
  first_name ILIKE 'BAR%';
```

## PostgreSQL Shorthand

| SQL Operator | PostgreSQL Shorthand |
|:------------:|:--------------------:|
| `LIKE`       | `~~`     |
| `NOT LIKE`   | `!~~`    |
| `ILIKE`      | `~~*`    |
| `NOT ILIKE`  | `!~~*`   |

Remembering from earlier look into docs, we can also use regex with `~` operator.

```{sql connection=con_dvd}
SELECT
  first_name,
  last_name
FROM
  customer
WHERE
  last_name ~ '^B.*e$';
```


## IS NULL

Connect to _**examples**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

First let's create a _contacts_ table.

```{sql connection=con_ex}
CREATE TABLE contacts(
  id          INT           GENERATED BY DEFAULT AS IDENTITY,
  first_name  VARCHAR(50)   NOT NULL,
  last_name   VARCHAR(50)   NOT NULL,
  email       VARCHAR(255)  NOT NULL,
  phone       VARCHAR(15),
  PRIMARY KEY(id)
);
```

Insert some records.

```{sql connection=con_ex}
INSERT INTO contacts
  (first_name, last_name, email, phone) 
VALUES
  ('John', 'Doe', 'john.doe@example.com', NULL),
  ('Lily', 'Bush', 'lily.bush@example.com', '(408-234-2764)');
```

```{sql connection=con_ex}
SELECT * FROM contacts;
```

`IS NULL` example:

```{sql connection=con_ex}
SELECT *
FROM
  contacts
WHERE
  phone IS NULL;
```

`IS NOT NULL` example:

```{sql connection=con_ex}
SELECT *
FROM
  contacts
WHERE
  phone IS NOT NULL;
```


## Alias

Alias template with `AS`:
```
SELECT
  column_name AS alias_name
FROM
  table_name;
```

Alias template w/o `AS` (because `AS` is optional):
```
SELECT
  column_name alias_name
FROM
  table_name;
```

Expression alias template:
```
SELECT 
  [expression] AS alias_name
FROM
  table_name;
```

_**Examples**_

The following query returns the full names of customers:

```{sql connection=con_dvd}
SELECT
  first_name || ' ' || last_name
FROM
  customer
ORDER BY
  first_name || ' ' || last_name;
```

We can use an alias to save some typing:

```{sql connection=con_dvd}
SELECT
  first_name || ' ' || last_name AS full_name
FROM
  customer
ORDER BY
  full_name;
```

Because PostgreSQL evaluates the `ORDER BY` clause after the `SELECT` clause, you can use the column alias in the `ORDER BY` clause.

For the other clauses evaluated before the `SELECT` clause such as `WHERE` `GROUP BY`, and `HAVING`, you cannot reference the column alias in these clauses.

### Table Alias

Table alias template:
```
SELECT
  column_1,
  column_2
FROM
  table_name AS alias_name;
```

These come in hand when tables have long names or when doing joins.

```
SELECT
  t1.column_name,
  t2.column_name
FROM
  table_name_1 AS t1
INNER JOIN
  table_name_2 AS t2
  ON join_predicate;
```

Finally, when doing a self-join, you _**must**_ use table aliases.

```
SELECT
  column_1,
  column_2
FROM
  table_name AS table_alias
INNER JOIN
  table_name ON table_alias.column_1 = table_name.column_2;
```

Quick example(s):

```{sql connection=con_dvd}
SELECT
  f.film_id,
  f.title,
  f.length,
  f.rating
FROM
  film AS f;
```


## Disconnect from DBs

```{r}
dbDisconnect(con_dvd)
dbDisconnect(con_ex)
```



```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
