---
title: "PostgreSQL Tutorial - Home - Section 4"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
---

# Joining Multiple Tables

## Joins

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```

Connect to _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

Set up sample tables.

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS basket_a (
  id     INT           PRIMARY KEY,
  fruit  VARCHAR(100)  NOT NULL
);
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS basket_b (
  id     INT           PRIMARY KEY,
  fruit  VARCHAR(100)  NOT NULL
);
```

```{sql connection=con_ex}
INSERT INTO basket_a (id, fruit)
VALUES
 (1, 'Apple'),
 (2, 'Orange'),
 (3, 'Banana'),
 (4, 'Cucumber');
```

```{sql connection=con_ex}
INSERT INTO basket_b (id, fruit)
VALUES
  (1, 'Orange'),
  (2, 'Apple'),
  (3, 'Watermelon'),
  (4, 'Pear');
```

```{sql connection=con_ex}
SELECT * FROM basket_a;
```

```{sql connection=con_ex}
SELECT * FROM basket_b;
```

### PostgreSQL inner join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
INNER JOIN
  basket_b AS b 
  ON a.fruit = b.fruit;
```

### PostgreSQL left outer join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
LEFT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit;
```

### PostgreSQL left outer join (where right table key is null)

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
LEFT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit
WHERE b.fruit IS NULL;
```

### PostgreSQL right outer join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
RIGHT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit;
```

### PostgreSQL right outer join (where left table key is null)

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
RIGHT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit
WHERE a.fruit IS NULL;
```

### PostgreSQL full outer join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
FULL OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit;
```

### PostgreSQL full outer join (where left and right table keys are null)

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
FULL OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit
WHERE a.id IS NULL OR b.id IS NULL;
```


## INNER JOIN














