---
title: "PostgreSQL Tutorial - Home - Section 8"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
    toc: true
    toc_float: true
---

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```


# Modifying data

## INSERT

When you create a new table, it doesn't have any data. You need to insert data into the table. `INSERT` allows you to insert one or more rows into a table at a time.

Here's the syntax for a single row insert.
```
INSERT INTO table_1
  (col_1, col_2, ...)
VALUES
  (value_1, value_2, ...);
```

Here's the syntax for a multiple row insert.
```
INSERT INTO table_1
  (col_1, col_2, ...)
VALUES
  (value_1a, value_2a, ...),
  (value_1b, value_2b, ...);
```

To insert data that comes form another table, you can use `INSERT INTO SELECT`.
```
INSERT INTO destination_table
  (d_col_1, d_col_2, ...)
SELECT
  s_col_A, 
  s_col_b, 
  ...
FROM
  source_table
WHERE          -- `WHERE` clause is optional
  [condition];
```

### `INSERT` examples

Let's create a table called "link".

Connect to the _**example**_ database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{sql connection=con_ex}
DROP TABLE IF EXISTS link;
```

```{sql connection=con_ex}
CREATE TABLE link(
  id           serial        PRIMARY KEY,
  url          VARCHAR(255)  NOT NULL,
  name         VARCHAR(255)  NOT NULL,
  description  VARCHAR(255),
  rel          VARCHAR(50)
);
```

Insert one row into the "link" table.

```{sql connection=con_ex}
INSERT INTO 
  link(url, name)
VALUES
  ('http://www.postgresqltutorial.com','PostgreSQL Tutorial');
```

Character data has to be enclosed in single quotes, `'`.

If you omit any column that accepts the `NULL` value in the `INSERT` statement, the column will take its default value. If not default value is set for the column, the column will take the `NULL` value.

Verify that the `INSERT` worked.

```{sql connection=con_ex}
SELECT * FROM link;
```

If you want to insert a string that contains a single quote character, such as _O'Reilly Media_, you have to use a single quote escape character.

```{sql connection=con_ex}
INSERT INTO 
  link(url, name)
VALUES
  ('http://www.oreilly.com', 'O''Reilly Media');
```

```{sql connection=con_ex}
SELECT * FROM link;
```

Insert multiple rows into the "link" table.

```{sql connection=con_ex}
INSERT INTO 
  link(url, name)
VALUES
  ('http://www.google.com','Google'),
  ('http://www.yahoo.com','Yahoo'),
  ('http://www.bing.com','Bing');
```

```{sql connection=con_ex}
SELECT * FROM link;
```

### Add a column with `ALTER TABLE`

Add a new column names `last_update` into the "link" table and set its default to `CURRENT_DATE`.

```{sql connection=con_ex}
ALTER TABLE 
  link
ADD COLUMN 
  last_update DATE;
```

```{sql connection=con_ex}
SELECT * FROM link;
```

```{sql connection=con_ex}
ALTER TABLE
  link
ALTER COLUMN
  last_update
SET DEFAULT CURRENT_DATE;
```

Insert a new row using the new `last_update` column.

```{sql connection=con_ex}
INSERT INTO
  link(url, name, last_update)
VALUES
  ('http://www.facebook.com', 'Facebook', '2013-06-01');
```

```{sql connection=con_ex}
SELECT * FROM link;
```

You can also use the `DEFAULT` keyword to set the default value for the `last_update` or any column that has a default value.

```{sql connection=con_ex}
INSERT INTO
  link(url, name, last_update)
VALUES
  ('https://www.tumblr.com/', 'Tumblr', DEFAULT);
```

```{sql connection=con_ex}
SELECT * FROM link;
```

What happens when I don't use `DEFAULT` in an `INSERT` statement?

```{sql connection=con_ex}
INSERT INTO
  link(url, name, description)
VALUES
  ('https://htdg.org', 'How to Design Programs', 'An introdution to programming and computing')
```

```{sql connection=con_ex}
SELECT * FROM link;
```

The answer is that (as explained above) the default value (`CURRENT_DATE`) is inserted into the field.

### Insert data from another table

```{sql connection=con_ex}
DROP TABLE IF EXISTS link_tmp;
```

```{sql connection=con_ex}
CREATE TABLE link_tmp (LIKE link);
```

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

```{sql connection=con_ex}
INSERT INTO
  link_tmp
SELECT
  *
FROM
  link
WHERE
  last_update IS NOT NULL;
```

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

### Get the last insert ID with PostgreSQL `RETURNING`

If you want to know the last value assigned to a serial field during an `INSERT`, use `RETURNING`.

```{sql connection=con_ex}
INSERT INTO
  link(url, name, last_update)
VALUES
  ('http://www.postgresql.org','PostgreSQL',DEFAULT)
RETURNING
  id;
```

This last `INSERT` shows that the `id` value is 9.

```{sql connection=con_ex}
SELECT * FROM link;
```

Disconnect from the _**example**_ database.
```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex); }
```



```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
