---
title: "PostgreSQL Tutorial - Home - Section 6"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
---

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```


# Performing set operations

## UNION 

The `UNION` operator combines result sets of two or more `SELECT` statements into a single result set. In a Venn diagram where set $A$ is the result of one `SELECT` statement and set $B$ is the result of another `SELECT` statement, `UNION` returns the union of A and B, $A \cup B$.

Here's the syntax.

```
SELECT
  column_1,
  column_2
FROM
  table_1
UNION
SELECT
  column_1,
  column_2
FROM
  table_2;
```

These rules apply to `UNION` queries:

1. Both queries must return the same number of columns.
2. The corresponding columns in the queries must have compatible data types.

The `UNION` operator removes all duplicate rows unless `UNION ALL` is used.

There are no guarantees when it comes to row ordering, so if you want a specific ordering, use `ORDER BY` on particular columns(s).

We often use the `UNION` operator to combine data from similar tables that are not perfectly normalized. Those tables are often found in the reporting system or data warehousing system.

Let's create some example tables.

Connect to the _**example**_ database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

Create the tables.
```{sql connection=con_ex}
DROP TABLE IF EXISTS sales2007q1;
DROP TABLE IF EXISTS sales2007q2;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS sales2007q1 (
  name    VARCHAR(64)  PRIMARY KEY,
  amount  MONEY        NOT NULL
);

CREATE TABLE IF NOT EXISTS sales2007q2 (
  name    VARCHAR(64)  PRIMARY KEY,
  amount  MONEY        NOT NULL
)
```

```{sql connection=con_ex}
INSERT INTO sales2007q1
  (name, amount)
VALUES
  ('Mike', 150000.25),
  ('Jon',  132000.75),
  ('Mary', 100000.00);
```

```{sql connection=con_ex}
INSERT INTO sales2007q2
  (name, amount)
VALUES
  ('Mike', 120000.25),
  ('Jon',  142000.75),
  ('Mary', 100000.00);
```

```{sql connection=con_ex}
SELECT * FROM sales2007q1;
```

```{sql connection=con_ex}
SELECT * FROM sales2007q2;
```

### `UNION` example

```{sql connection=con_ex}
SELECT
  *
FROM
  sales2007q1
UNION
SELECT
  *
FROM
  sales2007q2;
```

### `UNION ALL` example

```{sql connection=con_ex}
SELECT
  *
FROM
  sales2007q1
UNION ALL
SELECT
  *
FROM
  sales2007q2;
```

Here's a `UNION ALL` with an `ORDER BY`. (Note that the `ORDER BY` must go at the very end of `UNION [ALL]` query.)

```{sql connection=con_ex}
SELECT
  *
FROM
  sales2007q1
UNION ALL
SELECT
  *
FROM
  sales2007q2
ORDER BY
  name,
  amount DESC;
```


Do a little aggregate `SUM` with the `UNION ALL` result set as a subquery.

```{sql connection=con_ex}
SELECT
  name, SUM(amount)
FROM (
  SELECT
    *
  FROM
    sales2007q1
  UNION ALL
  SELECT
    *
  FROM
    sales2007q2
  ) AS sales2007q1q2
GROUP BY
  sales2007q1q2.name
ORDER BY
  SUM(amount) DESC;
```

Disconnect from the _**example**_ database.
```{r}
dbDisconnect(con_ex)
```



## INTERSECT

Like the `UNION` operator, the `INTERSECT` operator combines the result sets of two or more `SELECT` statements into a single result set. In a Venn diagram where set $A$ is the result of one `SELECT` statement and set $B$ is the result of another `SELECT` statement, `INTERSECT` returns the intersect of A and B, $A \cap B$.

The rules for an `INTERSECT` operator queries are the same as those for `UNION`:

1. Both queries must return the same number of columns.
2. The corresponding columns in the queries must have compatible data types.

Here's the syntax.

```
SELECT
  column_1,
  column_2
FROM
  table_1
INTERSECT
SELECT
  column_1,
  column_2
FROM
  table_2;
```

### `INTERSECT` operator example

Connect to the _**example**_ database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

Create the example tables.

```{sql connection=con_ex}
DROP TABLE IF EXISTS employees_2;
DROP TABLE IF EXISTS keys_2;
DROP TABLE IF EXISTS hipos_2;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS employees_2 (
  employee_id    SERIAL        PRIMARY KEY,
  employee_name  VARCHAR(255)  NOT NULL
);
```

```{sql connection=con_ex}
-- "key employees"
CREATE TABLE IF NOT EXISTS keys_2 (
  employee_id     INT   PRIMARY KEY,
  effective_date  DATE  NOT NULL --,
  -- FOREIGN KEY employee_id REFERENCES employees_2(employee_id)
);
```

```{sql connection=con_ex}
-- "high potential employees"
CREATE TABLE IF NOT EXISTS hipos_2 (
  employee_id     INT   PRIMARY KEY,
  effective_date  DATE  NOT NULL --,
  -- FOREIGN KEY employee_id REFERENCES employees_2(employee_id)
);
```

```{sql connection=con_ex}
INSERT INTO employees_2
  (employee_name)
VALUES
  ('Joyce Edwards'),
  ('Diane Collins'),
  ('Alice Stewart'),
  ('Julie Sanchez'),
  ('Heather Morris'),
  ('Teresa Rogers'),
  ('Doris Reed'),
  ('Gloria Cook'),
  ('Evelyn Morgan'),
  ('Jean Bell');

INSERT INTO keys_2
  (employee_id, effective_date)
VALUES
  (1, '2000-02-01'),
  (2, '2001-06-01'),
  (5, '2002-01-01'),
  (7, '2005-06-01');
 
INSERT INTO hipos_2
  (employee_id, effective_date)
VALUES
  (9, '2000-01-01'),
  (2, '2002-06-01'),
  (5, '2006-06-01'),
  (10, '2005-06-01');
```

```{sql connection=con_ex}
SELECT * FROM keys_2;
```

```{sql connection=con_ex}
SELECT * FROM hipos_2;
```

To find which employees are both "key" and "high impact", use the `INTERSECT` operator.

```{sql connection=con_ex}
SELECT
  employee_id
FROM
  keys_2
INTERSECT
SELECT
  employee_id
FROM
  hipos_2;
```

Here's an `INTERSECT` subquery with an `INNER JOIN` and an `ORDER BY`.

```{sql connection=con_ex}
SELECT
  keys_hipos.employee_id,
  employees_2.employee_name
FROM (
  SELECT
    employee_id
  FROM
    keys_2
  INTERSECT
  SELECT
    employee_id
  FROM
    hipos_2
) AS keys_hipos
INNER JOIN
  employees_2
  ON keys_hipos.employee_id = employees_2.employee_id
ORDER BY
  keys_hipos.employee_id DESC;
```

Disconnect from the _**example**_ database.
```{r}
dbDisconnect(con_ex)
```



```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
