---
title: "PostgreSQL Tutorial - Home - Section 4"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
---

# Joining Multiple Tables

## Joins

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```

Connect to _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

Set up sample tables.

```{sql connection=con_ex}
DROP TABLE IF EXISTS basket_a;
DROP TABLE IF EXISTS basket_b;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS basket_a (
  id     INT           PRIMARY KEY,
  fruit  VARCHAR(100)  NOT NULL
);

CREATE TABLE IF NOT EXISTS basket_b (
  id     INT           PRIMARY KEY,
  fruit  VARCHAR(100)  NOT NULL
);
```

```{sql connection=con_ex}
INSERT INTO basket_a (id, fruit)
VALUES
 (1, 'Apple'),
 (2, 'Orange'),
 (3, 'Banana'),
 (4, 'Cucumber');
```

```{sql connection=con_ex}
INSERT INTO basket_b (id, fruit)
VALUES
  (1, 'Orange'),
  (2, 'Apple'),
  (3, 'Watermelon'),
  (4, 'Pear');
```

```{sql connection=con_ex}
SELECT * FROM basket_a;
```

```{sql connection=con_ex}
SELECT * FROM basket_b;
```

### PostgreSQL inner join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
INNER JOIN
  basket_b AS b 
  ON a.fruit = b.fruit;
```

### PostgreSQL left outer join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
LEFT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit;
```

### PostgreSQL left outer join (where right table key is null)

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
LEFT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit
WHERE b.fruit IS NULL;
```

### PostgreSQL right outer join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
RIGHT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit;
```

### PostgreSQL right outer join (where left table key is null)

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
RIGHT OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit
WHERE a.fruit IS NULL;
```

### PostgreSQL full outer join

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
FULL OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit;
```

### PostgreSQL full outer join (where left and right table keys are null)

```{sql connection=con_ex}
SELECT
  a.id,
  a.fruit,
  b.id,
  b.fruit
FROM
  basket_a AS a
FULL OUTER JOIN
  basket_b AS b
  ON a.fruit = b.fruit
WHERE a.id IS NULL OR b.id IS NULL;
```

Disconnect from _**example**_ database.
```{r}
dbDisconnect(con_ex)
```


## INNER JOIN

Using the _**dvdrental**_ database...
```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

... we'll join the "customer" and "payment" tables.

```{sql connection=con_dvd}
SELECT
  c.customer_id,
  c.first_name,
  c.last_name,
  c.email,
  p.amount,
  p.payment_date
FROM
  customer AS c
INNER JOIN
  payment AS p
  ON c.customer_id = p.customer_id;
```

Add `ORDER BY` clause to sort by `customer_id` and `payment_date`.
```{sql connection=con_dvd}
SELECT
  c.customer_id,
  c.first_name,
  c.last_name,
  c.email,
  p.amount,
  p.payment_date
FROM
  customer AS c
INNER JOIN
  payment AS p
  ON c.customer_id = p.customer_id
ORDER BY
  c.customer_id, p.payment_date;
```

Add `WHERE` clause to filter.
```{sql connection=con_dvd}
SELECT
  c.customer_id,
  c.first_name,
  c.last_name,
  c.email,
  p.amount,
  p.payment_date
FROM
  customer AS c
INNER JOIN
  payment AS p
  ON c.customer_id = p.customer_id
WHERE
  c.customer_id = 2;
```

### INNER JOIN to join 3 tables

```{sql connection=con_dvd}
SELECT
  c.customer_id,
  c.first_name AS c_first_name,
  c.last_name AS c_last_name,
  c.email,
  s.first_name AS s_first_name,
  s.last_name AS s_last_name,
  p.amount,
  p.payment_date
FROM
  customer AS c
INNER JOIN
  payment AS p
  ON c.customer_id = p.customer_id
INNER JOIN
  staff AS s
  ON p.staff_id = s.staff_id;
```

Disconnect from _**dvdrental**_ database.
```{r}
dbDisconnect(con_dvd)
```


## LEFT JOIN

Connect to _**dvdrental**_ database.
```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

Left join "film" and "inventory" tables.
```{sql connection=con_dvd}
SELECT
  f.film_id,
  f.title,
  i.inventory_id
FROM
  film AS f
LEFT JOIN
  inventory AS i
  ON f.film_id = i.film_id;
-- WHERE i.inventory_id IS NULL; -- revealse right table NULLs
-- ORDER BY f.film_id;
```

Disconnect from _**dvdrental**_ database.
```{r}
dbDisconnect(con_dvd)
```


## SELF JOIN

Self-joins are useful for comparing rows within the same table.

Connect to the _**example**_ database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

### 1. Querying hierarchy data example

Set up a table based on the hierarchical reporting structure depicted in "PostgreSQL-Self-Join-Reporting-Structure.png".

Create an "employee" table.
```{sql connection=con_ex}
DROP TABLE IF EXISTS employee;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS employee(
  employee_id  INT           PRIMARY KEY,
  first_name   VARCHAR(255)  NOT NULL,
  last_name    VARCHAR(255)  NOT NULL,
  manager_id   INT,
  FOREIGN KEY (manager_id) REFERENCES employee (employee_id)
  ON DELETE CASCADE
);
```

Insert data into the "employee" table.
```{sql connection=con_ex}
INSERT INTO employee(
  employee_id, first_name, last_name, manager_id
) 
VALUES
  (1, 'Windy',  'Hays',       NULL),
  (2, 'Hasan',  'Conner',     1),
  (3, 'Ava',    'Christensn', 1),
  (4, 'Salley', 'Lester',     2),
  (5, 'Kelsie', 'Hays',       2),
  (6, 'Tory',   'Goff',       2),
  (7, 'Sau',    'Norman',     3),
  (8, 'Anna',   'Reeves',     3);
```

To find who reports to whom, use this INNER JOIN.
```{sql connection=con_ex}
SELECT
  e1.first_name || ' ' || e1.last_name AS employee_name,
  e2.first_name || ' ' || e2.last_name AS manager_name
FROM
  employee AS e1
INNER JOIN
  employee AS e2
  ON e1.manager_id = e2.employee_id
ORDER BY manager_name DESC;
```

A `LEFT OUTER JOIN` is required to include the top manager (Windy Hays) in result.
```{sql connection=con_ex}
SELECT
  e1.first_name || ' ' || e1.last_name AS employe_name,
  e2.first_name || ' ' || e2.last_name AS manager_name
FROM
  employee AS e1
LEFT OUTER JOIN
  employee AS e2
  ON e1.manager_id = e2.employee_id
ORDER BY
  manager_name DESC;
```

Disconnect from _**example**_ database.
```{r}
dbDisconnect(con_ex)
```


### 2. Comparing the rows with the same table

Connect to the _**dvdrental**_ database.
```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

The following query finds all pairs of films that have the same length.

```{sql connection=con_dvd}
SELECT
  f1.film_id,
  f1.title,
  f2.title,
  f1.length AS f1_length
FROM
  film AS f1
INNER JOIN
  film AS f2
  ON f1.length = f2.length;
-- ORDER BY
--   f1.film_id;
```

Disconnect from _**dvdrental**_ database.
```{r}
dbDisconnect(con_dvd)
```


## FULL OUTER JOIN

Connect to _**example**) database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

Create tables for this example.
```{sql connection=con_ex}
DROP TABLE IF EXISTS employees;
DROP TABLE IF EXISTS departments;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS departments(
  department_id    serial        PRIMARY KEY,
  department_name  VARCHAR(255)  NOT NULL
);
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS employees(
  employee_id    serial        PRIMARY KEY,
  employee_name  VARCHAR(255),
  department_id  INTEGER,
    FOREIGN KEY (department_id) REFERENCES departments(department_id)
);
```

Insert data into the tables.
```{sql connection=con_ex}
INSERT INTO departments 
  (department_name)
VALUES
  ('Sales'),
  ('Marketing'),
  ('HR'),
  ('IT'),
  ('Production');
```

```{sql connection=con_ex}
INSERT INTO employees
  (employee_name, department_id)
VALUES
  ('Bette Nicholson', 1),
  ('Christian Gable', 1),
  ('Joe Swank', 2),
  ('Fred Costner', 3),
  ('Sandra Kilmer', 4),
  ('Julia McQueen', NULL);
```

Query everything from both tables.
```{sql connection=con_ex}
SELECT * FROM departments;
```

```{sql connection=con_ex}
SELECT * FROM employees;
```

Use `FULL OUTER JOIN` to join all data from both tables.
```{sql connection=con_ex}
SELECT
  e.employee_id,
  e.employee_name,
  d.department_name
FROM
  employees AS e
FULL OUTER JOIN
  departments AS d
  ON e.department_id = d.department_id;
```

To find a department without any employees, use `WHERE`.
```{sql connection=con_ex}
SELECT
  e.employee_id,
  e.employee_name,
  d.department_name
FROM
  employees AS e
FULL OUTER JOIN
  departments AS d
  ON e.department_id = d.department_id
WHERE
  e.employee_id IS NULL;
```

Take a similar approach to find any employees not belonging to a department.
```{sql connection=con_ex}
SELECT
  e.employee_id,
  e.employee_name,
  d.department_id
FROM
  employees AS e
FULL OUTER JOIN
  departments AS d
  ON e.department_id = d.department_id
WHERE
  d.department_id IS NULL;
```

Close the _**example**_ database connection.
```{r}
dbDisconnect(con_ex)
```


## CROSS JOIN by Example

A `CROSS JOIN` clause allows you to produce the Cartesian Product of rows in two or more tables. Unlike the other joins, the `CROSS JOIN` doesn't have any matching condition in the join clause.

Suppose we have to perform the `CROSS JOIN` of two tables T1 and T2. For every row from T1 and T2 (i.e., a cartesian product), the result set will contain a table that consists of all columns in the T1 table followed by all columns in the T2 table. If T1 has $M$ rows, T2 has $N$ rows, the result set will have $N \times M$ rows.

Here's the syntax of a `CROSS JOIN`:
```
SELECT 
  *
FROM
  T1
CROSS JOIN
  T2;
```

Here's an alternative shorthand to the above:
```
SELECT 
  *
FROM 
  T1, T2;
```

The following is functionally equivalent to a `CROSS JOIN`:
```
SELECT
  *
FROM
  T1
INNER JOIN
  T2
  ON TRUE;
```

Let's create some example tables in the _**example**_ database.

First, connect to the _**example**_ database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

Create the example tables.
```{sql connection=con_ex}
DROP TABLE IF EXISTS T1;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS T1(
  label  CHAR(1)  PRIMARY KEY
);
```

```{sql connection=con_ex}
DROP TABLE IF EXISTS T2;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS T2(
  score  INT  PRIMARY KEY
);
```

```{sql connection=con_ex}
INSERT INTO T1
  (label)
VALUES
  ('A'),
  ('B');
```

```{sql connection=con_ex}
INSERT INTO T2
  (score)
VALUES
  (1),
  (2),
  (3);
```

Now for the `CROSS JOIN`:
```{sql connection=con_ex}
SELECT
  *
FROM
  T1
CROSS JOIN
  T2;
```

Let's try this out with tables that have 2 columns.

```{sql connection=con_ex}
DROP TABLE IF EXISTS T3;
DROP TABLE IF EXISTS T4;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS T3 (
  my_letter  CHAR(1)  PRIMARY KEY,
  my_number  INT      NOT NULL
);

CREATE TABLE IF NOT EXISTS T4 (
  my_char  CHAR(1)  PRIMARY KEY,
  my_int   INT      NOT NULL
);
```

```{sql connection=con_ex}
INSERT INTO T3
  (my_letter, my_number)
VALUES
  ('A', 1),
  ('B', 2),
  ('C', 3);
  
INSERT INTO T4
  (my_char, my_int)
VALUES
  ('C', 4),
  ('D', 5),
  ('E', 6);
```

```{sql connection=con_ex}
SELECT * FROM T3;
```

```{sql connection=con_ex}
SELECT * FROM T4;
```

Do the `CROSS JOIN`.
```{sql connection=con_ex}
SELECT
  *
FROM
  T3
CROSS JOIN
  T4;
```

Disconnect from the _**example**_ database.
```{r}
dbDisconnect(con_ex)
```


## NATURAL JOIN by Example

A natural join creates an implicit join based on the same column names in the joined tables. A natural join can be an `INNER`, `LEFT`, or `RIGHT`.

Here's the syntax.

```
SELECT
  *
FROM
  T1
NATURAL INNER JOIN
  T2;
```

```
SELECT
  *
FROM
  T1
NATURAL LEFT JOIN
  T2;
```

```
SELECT
  *
FROM
  T1
NATURAL RIGHT JOIN
  T2;
```

Let's create some example tables in the _**example**_ database.

First, connect to the _**example**_ database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{r}
dbListTables(con_ex)
```


Create the example tables `categories` and `products` and put some data in them.

```{sql connection=con_ex}
DROP TABLE IF EXISTS products;
DROP TABLE IF EXISTS categories;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS categories (
  category_id    serial        PRIMARY KEY,
  category_name  VARCHAR(255)  NOT NULL
);

CREATE TABLE IF NOT EXISTS products (
  product_id    serial        PRIMARY KEY,
  product_name  VARCHAR(255)  NOT NULL,
  category_id   INT           NOT NULL,
  FOREIGN KEY (category_id) REFERENCES categories (category_id)
);
```

```{sql connection=con_ex}
INSERT INTO categories 
  (category_name)
VALUES
  ('Smart Phone'),
  ('Laptop'),
  ('Tablet');
  
INSERT INTO products
  (product_name, category_id)
VALUES
  ('iPhone',          1),
  ('Samsung Galaxy',  1),
  ('HP Elite',        2),
  ('Lenovo Thinkpad', 2),
  ('iPad',            3),
  ('Kindle Fire',     3);
```

```{sql connection=con_ex}
SELECT * FROM categories;
```

```{sql connection=con_ex}
SELECT * FROM products;
```

Here's the result of a `NATURAL INNER JOIN`.

```{sql connection=con_ex}
SELECT
  *
FROM
  products
NATURAL INNER JOIN
  categories;
```

The above is the same as this explicit `INNER JOIN`.

```{sql connection=con_ex}
SELECT
  products.category_id,
  products.product_id,
  products.product_name,
  categories.category_name
FROM
  products
INNER JOIN
  categories
  ON products.category_id = categories.category_id;
```

The advantage is that the `NATURAL JOIN` doesn't require that we explicitly specify the `ON ... = ...` clause. As long as the fields that represent the same data have matching names, the `NATURAL JOIN` implicitly joins them.

However, the convenience of this implicit joining can cause unexpected results if we're not careful.

Disconnect form the _**example**_ database.

```{r}
dbDisconnect(con_ex)
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
