---
title: "PostgreSQL Tutorial - Home - Section 5"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
---

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```


# Grouping data

## GROUP BY

Connect to _**dvdrental**_ database.

```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

`GROUP BY` using aggregate function template.
```
SELECT
  column_1, aggregate_function(column_2)
FROM
  table_name
GROUP_BY
  column_1;
```

The `GROUP BY` clause must appear right after the `FROM` or `WHERE` clause. The `GROUP BY` is one column or a comma-separated list of columns. You can also put an expression in the `GROUP BY` clause.

### `GROUP BY` without aggregate function

Here, `GROUP BY` acts like the `DISTINCT` clause... removing duplicates from the result set.

```{sql connection=con_dvd}
SELECT
  customer_id
FROM
  payment
GROUP BY
  customer_id;
```

### `GROUP BY` with `SUM` aggregate function

```{sql connection=con_dvd}
SELECT
  customer_id,
  SUM(amount)
FROM
  payment
GROUP BY
  customer_id;
```

```{sql connection=con_dvd}
SELECT
  customer_id,
  SUM(amount)
FROM
  payment
GROUP BY
  customer_id
ORDER BY
  SUM(amount) DESC;
```

### `GROUP BY` with `COUNT` aggregate function 

```{sql connection=con_dvd}
SELECT
  staff_id,
  COUNT(payment_id)
FROM
  payment
GROUP BY
  staff_id;
```

Disconnect from _**dvdrental**_ database.
```{r}
dbDisconnect(con_dvd)
```


## HAVING

`HAVING` is used in conjunction with `GROUP BY` to filter group rows that do not satisfy some condition.

Here's the syntax.
```
SELECT
  column_1,
  aggregate_function(column_2)
FROM
  table_name
GROUP BY
  column_1
HAVING
  aggregate_function(column_2) > 2.00; -- or whatever condition
```

`HAVING` is like `WHERE`, but `WHERE` is used for conditions _**BEFORE**_ the grouping, and `HAVING` is used for conditions _**AFTER**_ the grouping.

Here are some examples.

Connect to _**dvdrental**_ database.

```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

### `HAVING` clause with `SUM` function example

```{sql connection=con_dvd}
SELECT
  customer_id,
  SUM(amount)
FROM
  payment
GROUP BY
  customer_id
ORDER BY
  SUM(amount) DESC;
```

```{sql connection=con_dvd}
SELECT
  customer_id,
  SUM(amount)
FROM
  payment
GROUP BY
  customer_id
HAVING
  SUM(amount) > 200.00
ORDER BY
  SUM(amount) DESC;
```

### `HAVING` clause with `COUNT` and `AS` example

```{sql connection=con_dvd}
SELECT
  store_id,
  COUNT(customer_id) AS customer_count
FROM
  customer
GROUP BY
  store_id
ORDER BY
 customer_count DESC;
```

```{sql connection=con_dvd}
SELECT
  store_id,
  COUNT(customer_id) AS customer_count
FROM
  customer
GROUP BY
  store_id
HAVING
  COUNT(customer_id) > 300 -- aggregate function in HAVING clause
ORDER BY
  customer_count DESC;     -- alias in ORDER BY clause
```

### `HAVING` without `GROUP BY` clause

In PostgreSQL, you can use the `HAVING` clause without the `GROUP BY` clause. In this case, the `HAVING` clause will turn the query into a single group. In addition, the `SELECT` list and `HAVING` clause can only refer to columns from within aggregate functions. This kind of query returns a single row if the condition in the `HAVING` clause is true or zero row if it is false.

```{sql connection=con_dvd}
SELECT
  COUNT(customer_id)
FROM
  customer
HAVING
  COUNT(customer_id) > 300;
```

```{sql connection=con_dvd}
SELECT
  COUNT(customer_id)
FROM
  customer
HAVING
  COUNT(customer_id) > 600;
```

Disconnect from _**dvdrental**_ database.
```{r}
dbDisconnect(con_dvd)
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
