---
title: "PostgreSQL Tutorial - Home - Section 7"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
---

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```


# Grouping Sets

## GROUPING SETS 

Let's start by creating an example table called "sales".

Connect to the _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{sql connection=con_ex}
DROP TABLE IF EXISTS sales;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS sales (
  brand     VARCHAR  NOT NULL,
  segment   VARCHAR  NOT NULL,
  quantity  INT      NOT NULL,
  PRIMARY KEY (brand, segment)
);
```

```{sql connection=con_ex}
INSERT INTO sales
  (brand, segment, quantity)
VALUES
  ('ABC', 'Premium', 100),
  ('ABC', 'Basic',   200),
  ('XYZ', 'Premium', 100),
  ('XYZ', 'Basic',   300);
```

```{sql connection=con_ex}
SELECT * FROM sales;
```

### `GROUPING SETS` example

A grouping set is a set of columns by which you group. Typically, a single aggregate query defines a single grouping set.

For example, this query defines a grouping set of the `brand` and `segment`. It returns the number of products sold by `brand` and `segment`.

```{sql connection=con_ex}
SELECT
  brand,
  segment,
  SUM(quantity)
FROM
  sales
GROUP BY
  brand,
  segment;
```

This query returns the number of products sold by `brand`. It defines a grouping set of the `brand`.

```{sql connection=con_ex}
SELECT
  brand,
  SUM(quantity)
FROM
  sales
GROUP BY
  brand;
```

This query returns the number of products sold by `segment`. It defines a grouping set of the `segment`.

```{sql connection=con_ex}
SELECT
  segment,
  SUM(quantity)
FROM
  sales
GROUP BY
  segment;
```

This query returns the number of products sold for all `brand`s and `segment`s. It defines an empty grouping set.

```{sql connection=con_ex}
SELECT
  SUM(quantity)
FROM
  sales;
```

Suppose instead of the four separate result sets above, you wanted to see a unified result set with the aggregated data for all grouping sets. To achieve this, you use the `UNION ALL` to unify all the queries above.

Because the `UNION ALL` requires all result sets to have the same number of columns with compatible data types, you need to adjust the queries by adding `NULL` to the selection list of each as shown below.

```{sql connection=con_ex}
SELECT
  brand,
  segment,
  SUM(quantity)
FROM
  sales
GROUP BY
  brand,
  segment

UNION ALL

SELECT
  brand,
  NULL,
  SUM(quantity)
FROM
  sales
GROUP BY
  brand

UNION ALL

SELECT
  NULL,
  segment,
  SUM(quantity)
FROM
  sales
GROUP BY
  segment

UNION ALL

SELECT
  NULL,
  NULL,
  SUM(quantity)
FROM
  sales;
```

This query with the `UNION ALL`s has two disadvantages:

1. It's super long to write.
2. It has performance issues because the PostgreSQL engine has to scan the "sales" table separately for each of the four (sub)queries joined by `UNION ALL`s.

PostgreSQL provides `GROUPING SETS` -- which is a subclause of `GROUP BY` -- to make this kind of query more efficient.

Here's the syntax.

```
SELECT
  column_1,
  column_2,
  aggregate_function(column_3)
FROM
  table_name
GROUP BY
  GROUPING SETS (
    (column_1, column_2),
    (column_1),
    (column_2),
    ()
  );
```

Applying this syntax to your previous `UNION ALL` approach, we get this.

```{sql connection=con_ex}
SELECT
  brand,
  segment,
  SUM(quantity)
FROM
  sales
GROUP BY
  GROUPING SETS(
    (brand, segment),
    (brand),
    (segment),
    ()
  );
```

Using `GROUPING SETS` is much easier to read and the PostgreSQL engine optimizes the return speed by reducing the number of scans it has to make of the "sales" table.


## GROUPING function

The `GROUPING` function accepts a name of a column and returns bit 0 if the row is a member of the current grouping set and 1 otherwise.

### `GROUPING` function example

Here's an example.

```{sql connection=con_ex}
SELECT
  GROUPING(brand) AS grouping_brand,
  GROUPING(segment) AS grouping_segment,
  brand,
  segment,
  SUM(quantity)
FROM
  sales
GROUP BY
  GROUPING SETS(
    (brand, segment),
    (brand),
    (segment),
    ()
  )
ORDER BY
  grouping_brand,
  grouping_segment,
  brand,
  segment;
```

0 + 0 means the row is in the `(brand, segment)` grouping set.

0 + 1 means the row is in the `(brand)` grouping set.

1 + 0 means the row is in the `(segment)` grouping set.

1 + 1 means the row is in the `()` grouping set.



Disconnect from the _**example**_ database.

```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex) }
```


## CUBE

`CUBE` is a shorthand for listing all possible grouping sets from a list of columns.

For example if we want to group by `col_1`, `col_2`, and `col_3`, all possible grouping sets are `(col_1, col_2, col_3)`, `(col_1, col_2)`, `(col_1, col_3)`, `(col_2, col_3)`, `(col_1)`, `(col_2)`, `(col_3)`, and `()`.

So instead of writing this...
```
SELECT
  col_1,
  col_2,
  col_3,
  aggregate_function(col_4)
FROM
  table_1
GROUP BY
  GROUPING SETS (
    (col_1, col_2, col_3),
    (col_1, col_2),
    (col_1, col_3),
    (col_2, col_3),
    (col_1),
    (col_2),
    (col_3),
    ()
  );  
```

... we can write this...

```
SELECT
  col_1,
  col_2,
  col_3,
  aggregate_function(col_4)
FROM
  table_1
GROUP BY
  CUBE (
    col_1,
    col_2,
    col_3
  );
```

You can also do a partial `CUBE` grouping.

```
SELECT
  col_1,
  col_2,
  col_3,
  aggregate_function(col_4)
FROM
  table_1
GROUP BY
  col_1,
  CUBE (
    col_1,
    col_2
  );
```

### `CUBE` example

Connect to the _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

Let's use the "sales" table we already created.

```{sql connection=con_ex}
SELECT * FROM sales;
```

```{sql connection=con_ex}
SELECT
  brand,
  segment,
  SUM(quantity),
  GROUPING(brand) AS grouping_brand,
  GROUPING(segment) AS grouping_segment
FROM
  sales
GROUP BY
  CUBE(
    brand,
    segment
  )
ORDER BY
  grouping_brand,
  grouping_segment,
  brand,
  segment;
```

This is a partial cube.

```{sql connection=con_ex}
SELECT
  brand,
  segment,
  SUM(quantity),
  GROUPING(brand) AS grouping_brand,
  GROUPING(segment) AS grouping_segment
FROM
  sales
GROUP BY
  brand,
  CUBE(segment) -- (segment) & () => (brand, segment) & (brand)
ORDER BY
  grouping_brand,
  grouping_segment,
  brand,
  segment;
```

Disconnect from _**example**_ database.

```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex) }
```


## ROLLUP

Different from the `CUBE` subclause, `ROLLUP` does _**not**_ generate all possible grouping sets based on the specified columns. `ROLLUP` only generates a subset of all possible grouping sets based on the assumption of a column grouping hierarchy. 
For example, using `col_1`, `col_2`, and `col_3` as grouping columns, `CUBE` generates all possible grouping sets:

```
(col_1, col_2, col_3)
(col_1, col_2)
(col_1, col_3)
(col_2, col_3)
(col_1)
(col_2)
(col_3)
()
```

But the subclause `ROLLUP (col_1, col_2, col_3)` only generates four grouping subsets assuming the hierarchy `col_1` $>$ `col_2` $>$ `col_3` as follows:

```
(col_1, col_2, col_3)
(col_1, col_2)
(col_1)
()
```

A common use of `ROLLUP` is to calculate aggregations by year, month, and date (assuming the hierarchy year $>$ month $>$ date).

Here's the syntax.

```
SELECT
  col_1,
  col_2,
  col_3,
  aggregate_function(col_4)
FROM
  table_1
GROUP BY
  ROLLUP (
    col_1,
    col_2,
    col_3
  );
```

It's also possible to partial rollups to reduce the number of subtotals generated.

```
SELECT
  col_1,
  col_2,
  col_3,
  aggregate_function(col_4)
FROM
  table_1
GROUP BY
  col_1
  ROLLUP (
    col_2,
    col_3
  );
```

### `ROLLUP` example

Let's use the previously used "sales" table as an example.

Connect to the _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{sql connection=con_ex}
SELECT
  brand,
  segment,
  SUM(quantity),
  GROUPING(brand) AS grouping_brand,
  GROUPING(segment) AS grouping_segment
FROM
  sales
GROUP BY
  ROLLUP(
    brand,
    segment
  ) -- => (brand, segment), (brand), ()
ORDER BY
  grouping_brand,
  grouping_segment,
  brand,
  segment;
```

Here's what happens when we change the column grouping hierarchy by swapping the order of `brand` and `segment`.

```{sql connection=con_ex}
SELECT
  segment,
  brand,
  SUM(quantity),
  GROUPING(segment) AS grouping_segment,
  GROUPING(brand) AS grouping_brand
FROM
  sales
GROUP BY
  ROLLUP(
    segment,
    brand
  ) -- => (segment, brand), (segment), ()
ORDER BY
  grouping_segment,
  grouping_brand,
  segment,
  brand;
```

And here's a partial `ROLLUP` example.

```{sql connection=con_ex}
SELECT
  brand,
  segment,
  SUM(quantity),
  GROUPING(brand) as grouping_brand,
  GROUPING(segment) as grouping_segment
FROM
  sales
GROUP BY
  brand,
  ROLLUP(
    segment
  ) -- => (brand, segment), (brand)
ORDER BY
  grouping_brand,
  grouping_segment,
  brand,
  segment;
```

And here's an example using the _**dvdrental**_ database.

```{r}
con_dvd <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```

Let's use the "rental" table.

First, let's preview the table.

```{sql connection=con_dvd}
SELECT * FROM rental LIMIT 5;
```

We can use `COUNT` on the `rental_id` column and extract the year, month, and day from the `rental_date` column.

```{sql connection=con_dvd}
SELECT
  EXTRACT(YEAR FROM rental_date) AS year,
  EXTRACT(MONTH FROM rental_date) AS month,
  EXTRACT(DAY FROM rental_date) AS day,
  COUNT(rental_id) --,
  -- GROUPING(EXTRACT(YEAR FROM rental_date)) AS grouping_year,
  -- GROUPING(EXTRACT(MONTH FROM rental_date)) AS grouping_month,
  -- GROUPING(EXTRACT(DAY FROM rental_date)) AS grouping_day
FROM
  rental
GROUP BY
  ROLLUP(
    year,
    month,
    day
  )
ORDER BY
  -- grouping_year,
  -- grouping_month,
  -- grouping_day,
  year,
  month,
  day;
```

Disconnect from both the _**example**_ and _**dvdrental**_ databases.

```{r}
if(exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex) }
if(exists("con_dvd")) { dbDisconnect(con_dvd); rm(con_dvd) }
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
