---
title: "PostgreSQL Tutorial - Home - Section 8"
output: 
  html_notebook:
    theme: cerulean
    highlight: espresso
    toc: true
    toc_float: true
---

Load libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```


# Modifying data

## INSERT

When you create a new table, it doesn't have any data. You need to insert data into the table. `INSERT` allows you to insert one or more rows into a table at a time.

Here's the syntax for a single row insert.
```
INSERT INTO table_1
  (col_1, col_2, ...)
VALUES
  (value_1, value_2, ...);
```

Here's the syntax for a multiple row insert.
```
INSERT INTO table_1
  (col_1, col_2, ...)
VALUES
  (value_1a, value_2a, ...),
  (value_1b, value_2b, ...);
```

To insert data that comes form another table, you can use `INSERT INTO SELECT`.
```
INSERT INTO destination_table
  (d_col_1, d_col_2, ...)
SELECT
  s_col_A, 
  s_col_b, 
  ...
FROM
  source_table
WHERE          -- `WHERE` clause is optional
  [condition];
```

### `INSERT` examples

Let's create a table called "link".

Connect to the _**example**_ database.
```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{sql connection=con_ex}
DROP TABLE IF EXISTS link;
```

```{sql connection=con_ex}
CREATE TABLE link(
  id           serial        PRIMARY KEY,
  url          VARCHAR(255)  NOT NULL,
  name         VARCHAR(255)  NOT NULL,
  description  VARCHAR(255),
  rel          VARCHAR(50)
);
```

Insert one row into the "link" table.

```{sql connection=con_ex}
INSERT INTO 
  link(url, name)
VALUES
  ('http://www.postgresqltutorial.com','PostgreSQL Tutorial');
```

Character data has to be enclosed in single quotes, `'`.

If you omit any column that accepts the `NULL` value in the `INSERT` statement, the column will take its default value. If not default value is set for the column, the column will take the `NULL` value.

Verify that the `INSERT` worked.

```{sql connection=con_ex}
SELECT * FROM link;
```

If you want to insert a string that contains a single quote character, such as _O'Reilly Media_, you have to use a single quote escape character.

```{sql connection=con_ex}
INSERT INTO 
  link(url, name)
VALUES
  ('http://www.oreilly.com', 'O''Reilly Media');
```

```{sql connection=con_ex}
SELECT * FROM link;
```

Insert multiple rows into the "link" table.

```{sql connection=con_ex}
INSERT INTO 
  link(url, name)
VALUES
  ('http://www.google.com','Google'),
  ('http://www.yahoo.com','Yahoo'),
  ('http://www.bing.com','Bing');
```

```{sql connection=con_ex}
SELECT * FROM link;
```

### Add a column with `ALTER TABLE`

Add a new column names `last_update` into the "link" table and set its default to `CURRENT_DATE`.

```{sql connection=con_ex}
ALTER TABLE 
  link
ADD COLUMN 
  last_update DATE;
```

```{sql connection=con_ex}
SELECT * FROM link;
```

```{sql connection=con_ex}
ALTER TABLE
  link
ALTER COLUMN
  last_update
SET DEFAULT CURRENT_DATE;
```

Insert a new row using the new `last_update` column.

```{sql connection=con_ex}
INSERT INTO
  link(url, name, last_update)
VALUES
  ('http://www.facebook.com', 'Facebook', '2013-06-01');
```

```{sql connection=con_ex}
SELECT * FROM link;
```

You can also use the `DEFAULT` keyword to set the default value for the `last_update` or any column that has a default value.

```{sql connection=con_ex}
INSERT INTO
  link(url, name, last_update)
VALUES
  ('https://www.tumblr.com/', 'Tumblr', DEFAULT);
```

```{sql connection=con_ex}
SELECT * FROM link;
```

What happens when I don't use `DEFAULT` in an `INSERT` statement?

```{sql connection=con_ex}
INSERT INTO
  link(url, name, description)
VALUES
  ('https://htdg.org', 'How to Design Programs', 'An introdution to programming and computing')
```

```{sql connection=con_ex}
SELECT * FROM link;
```

The answer is that (as explained above) the default value (`CURRENT_DATE`) is inserted into the field.

### Insert data from another table

```{sql connection=con_ex}
DROP TABLE IF EXISTS link_tmp;
```

```{sql connection=con_ex}
CREATE TABLE link_tmp (LIKE link);
```

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

```{sql connection=con_ex}
INSERT INTO
  link_tmp
SELECT
  *
FROM
  link
WHERE
  last_update IS NOT NULL;
```

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

### Get the last insert ID with PostgreSQL `RETURNING`

If you want to know the last value assigned to a serial field during an `INSERT`, use `RETURNING`.

```{sql connection=con_ex}
INSERT INTO
  link(url, name, last_update)
VALUES
  ('http://www.postgresql.org','PostgreSQL',DEFAULT)
RETURNING
  id;
```

This last `INSERT` shows that the `id` value is 9.

```{sql connection=con_ex}
SELECT * FROM link;
```

Disconnect from the _**example**_ database.
```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex); }
```


## UPDATE

To change the values of columns in a table, use `UPDATE`.

Here's the syntax.

```
UPDATE 
  table_1
SET
  col_1 = value_1,
  col_2 = value_2,
  ...
WHERE
  [condition];
```

If you omit the `WHERE` clause, all the rows in a column will be updated.

### `UPDATE` examples

We'll use the "link" table created above.

Connect to the _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{sql connection=con_ex}
SELECT * FROM link;
```

Let's change the `NULL` values in the `last_update` column to the `DEFAULT` value (i.e., `CURRENT_DATE`).

```{sql connection=con_ex}
UPDATE
  link
SET
  last_update = DEFAULT
WHERE
  last_update IS NULL;
```

```{sql connection=con_ex}
SELECT * FROM link;
```

Now let's update all rows in a particular column. To do this, omit the `WHERE` clause.

```{sql connection=con_ex}
UPDATE
  link
SET
  rel = 'nofollow';
```

```{sql connection=con_ex}
SELECT * FROM link;
```

Update a column from another column within the same table.

```{sql connection=con_ex}
UPDATE
  link
SET
  description = name;
```

```{sql connection=con_ex}
SELECT * FROM link;
```

### `UPDATE` join example

Let's look at the "link_tmp" table.

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

Here we can update corresponding values in the "link_tmp" table that come from "link" table.

```{sql connection=con_ex}
UPDATE
  link_tmp
SET
  rel = link.rel,
  description = link.description,
  last_update = link.last_update
FROM
  link
WHERE
  link_tmp.id = link.id;
```

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

### `UPDATE` with `RETURNING` example

The `UPDATE` statement returns the number of affected rows by default. The PostgreSQL `UPDATE` statement also returns updated entries using the `RETURNING` clause. This addition is a PostgreSQL’s extension to the SQL standard.

```{sql connection=con_ex}
UPDATE
  link
SET
  description = 'Learn PostgreSQL fast and easy',
  rel = 'follow'
WHERE
  id = 1
RETURNING
  id,
  description,
  rel;
```

Verify the udpate.

```{sql connection=con_ex}
SELECT 
  * 
FROM 
  link
WHERE
  id = 1;
```

Disconnect from the _**example**_ database.

```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex); }
```


## UPDATE Join

Sometimes you need to update data of a table based on values in another table. You can use the "`UPDATE` join" in PostgreSQL to do this.

Here's the syntax.

```
UPDATE
  table_A
SET
  A.c1 = expression
FROM
  table_B
WHERE
  A.c2 = B.c2;
```

Notice that the `FROM` clause must appear immediately after the `SET` clause.

We'll create a couple new tables for examples.

Connect to the _**example**_ database

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{r}
dbListTables(con_ex)
```

```{sql connection=con_ex}
DROP TABLE IF EXISTS product_segment;
DROP TABLE IF EXISTS product;
```

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS product_segment (
  id        SERIAL   PRIMARY KEY,
  segment   VARCHAR  NOT NULL,
  discount  NUMERIC  (4, 2)
);
```

The "product_segment" table has the discount column that stores the discount percentage based on a specific segment. For example, grand luxury products have 5% discount while luxury and mass products have 6% and 10% discount respectively.

```{sql connection=con_ex}
INSERT INTO
  product_segment(segment, discount)
VALUES
  ('Grand Luxury', 0.05),
  ('Luxury', 0.06),
  ('Mass', 0.10);
```

```{sql connection=con_ex}
SELECT * FROM product_segment;
```

Second, create another table named "product" that stores the product data. The product table has the foreign key column `segment_id` that links to the id of the segment table.

```{sql connection=con_ex}
CREATE TABLE IF NOT EXISTS product(
  id SERIAL PRIMARY KEY,
  name varchar NOT NULL,
  price NUMERIC(10, 2),
  net_price NUMERIC(10, 2),
  segment_id INT NOT NULL,
  FOREIGN KEY (segment_id) REFERENCES product_segment(id)
);
```

```{sql connection=con_ex}
INSERT INTO 
  product(name, price, segment_id) 
VALUES 
  ('diam',                    804.89, 1),
  ('vestibulum aliquet',      228.55, 3),
  ('lacinia erat',            366.45, 2),
  ('scelerisque quam turpis', 145.33, 3),
  ('justo lacinia',           551.77, 2),
  ('ultrices mattis odio',    261.58, 3),
  ('hendrerit',               519.62, 2),
  ('in hac habitasse',        843.31, 1),
  ('orci eget orci',          254.18, 3),
  ('pellentesque',            427.78, 2),
  ('sit amet nunc',           936.29, 1),
  ('sed vestibulum',          910.34, 1),
  ('turpis eget',             208.33, 3),
  ('cursus vestibulum',       985.45, 1),
  ('orci nullam',             841.26, 1),
  ('est quam pharetra',       896.38, 1),
  ('posuere',                 575.74, 2),
  ('ligula',                  530.64, 2),
  ('convallis',               892.43, 1),
  ('nulla elit ac',           161.71, 3);
```

```{sql connection=con_ex}
SELECT * FROM product;
```

Suppose you want to calculate the net price of every product based on the discount of the product segment. To do this, do an "`UPDATE` join".

```{sql connection=con_ex}
UPDATE
  product
SET
  net_price = price - (price * discount)
FROM
  product_segment
WHERE
  product.segment_id = product_segment.id;
```

```{sql connection=con_ex}
SELECT * FROM product;
```

This is cool!

Disconnect from the _**example**_ database.

```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex); }
```


## UPSERT





## DELETE

To delete data from a table, use `DELETE`.

Here's the syntax.

```
DELETE FROM
  table_1
WHERE
  [condition];
```

*If you omit the `WHERE` clause, the `DELETE` statement will DELETE ALL THE ROWS FROM THE TABLE!*

The `DELETE` statement returns the number of deleted rows. It may not the same as the number of rows specified by the condition in the `WHERE` clause because the table may contain a `BEFORE DELETE` trigger that does something else before deletion. If no row is deleted, the `DELETE` statement returns zero.

If you want to check the condition that references one or more columns from another table, use the `USING` clause.

```
DELETE FROM
  table_1
USING
  table_2
WHERE
  table_1.id = table_2.id
  AND ...;
```

You can also use a subquery to do the same thing.

```
DELETE FROM
  table_1
WHERE table_1.id = (SELECT id from table_2);
```

### `DELETE` examples

We will use the "link" and "link_tmp" tables that we created.

Connect to the _**example**_ database.

```{r}
con_ex <- dbConnect(odbc::odbc(), "PostgreSQL AWS example")
```

```{sql connection=con_ex}
SELECT * FROM link;
```

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

`DELETE` using the `WHERE` clause.

```{sql connection=con_ex}
DELETE FROM
  link
WHERE
  id = 8;
```

```{sql connection=con_ex}
SELECT * FROM link;
```

`DELETE` using a `USING` clause... suppose you want to remove all the rows in the "link" table that have ID values in `id` column of the "link_tmp" table (i.e., `6`, `7`, and `8`).

```{sql connection=con_ex}
DELETE FROM
  link
USING
  link_tmp
WHERE
  link.id = link_tmp.id;
```

```{sql connection=con_ex}
SELECT * FROM link;
```

`DELETE` all the rows from a table.

```{sql connection=con_ex}
DELETE FROM
  link;
```

```{sql connection=con_ex}
SELECT * FROM link;
```

To delete all the rows in the "link_tmp" table and return the deleted rows, use the `RETURNING` clause in the `DELETE` statement.

```{sql connection=con_ex}
DELETE FROM
  link_tmp
RETURNING
  *; -- return all columns
```

```{sql connection=con_ex}
SELECT * FROM link_tmp;
```

Disconnect from the _**example**_ database.

```{r}
if (exists("con_ex")) { dbDisconnect(con_ex); rm(con_ex); }
```




```{r echo=FALSE}
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
